---
title: "LAB4 Multiple Linear Regression"
author: "Samuel Mweni"
date: "Stat 415/615 Regression, `r lubridate::year(Sys.Date())`"
output:
  word_document:
    toc: yes
    toc_depth: '1'
  pdf_document:
    toc: yes
    toc_depth: 1
    number_sections: yes
urlcolor: blue
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align  = "center")
```

1  Commercial Property (polynomial and interactions)
```{r}
cpdata<-read.table("../homework7/CommercialProperty.txt", header=T)
#cpdata
head(cpdata, 3) # first 3 observations
```
```{r}
tail(cpdata, 3) # last 3 observations
```
#1.1 Plot the data and comment. Also, get a numerical summary (mean,std.dev.) of all variables. Note that the mean of Age is 7.86.

```{r}
par(mfrow=c(2,3))
for (i in 1:5) hist(cpdata[, i])
summary(cpdata)
```
```{r}
pairs(cpdata)
```

```{r}
cor(cpdata)

```
#1.2 Follow the instruction in Problem 8.8 from the text, we will consider a regression model with predictors Age, Ageˆ2, Expense, and Square Footage.Fit the regression model.
```{r}
cpdata$CentAge<-cpdata$Age-mean(cpdata$Age)
cpreg1 <- lm(RentalRates~Age+I(Age^2)+Expense+Sfootage, data=cpdata)
summary(cpreg1)
```
#1.3 Add interaction terms to the previous model and examine their signifiance.

```{r}
cpreg2<-lm(RentalRates~Age+I(Age^2)+Expense+Sfootage
+ Age:Expense + Age:Sfootage, data=cpdata)
summary(cpreg2)
```

```{r}
# Using centered-variables.
cpreg2b<-lm(RentalRates~CentAge+I(CentAge^2)+Expense+Sfootage
+ CentAge:Expense + CentAge:Sfootage, data=cpdata)
cpreg2b

```
```{r}
anova(cpreg1, cpreg2)

```
```{r}
anova(cpreg1, cpreg2)

```

# 2 Example 2. Qualitative/Categorical Predictors and Interactions
The data in file twins.txt and twins.sav are from a 1966 paper by Cyril Burt entitled “The genetic
determination of differences in intelligence: A study of monozygotic twins reared apart”. The data consist of
IQ scores for identical twins, one raised by foster parents, the other by the natural parents. We also know the
social class of natural parents (high, middle or low). We are interested in predicting the IQ of the twin with
foster parents from the IQ of the twin with the natural parents and the social class of natural parents.
```{r}
twindata<-read.table("../homework7/twins.txt", header=T)
head(twindata, 2)

```
```{r}
tail(twindata, 2)

```

```{r}
# Covert status to factor to simplify future code.
twindata$status <- as.factor(twindata$status)

```
2.1 Plot the data
```{r}
par(mfrow = c(1, 3))
hist(twindata$IQF)
hist(twindata$IQN)
barplot(table(twindata$status))
```
```{r}
plot(IQF ~ IQN, col=status, pch=14+as.numeric(status), data=twindata)
legend(70, 130, legend=c("High", "Low", "Median"), col=c(1:3), pch=14+c(1:3))
```
# 2.3 Fit regression models.
Recall that we converted the categorical variable status into a factor variable earlier. Hence it can be
used in the lm() function directly. If a categorical predictor is not converted into a factor, you must use
as.factor(status) inside the lm() function.


```{r}
twinreg<-lm(IQF~IQN+status, data=twindata)
# twinreg<-lm(IQF~IQN+as.factor(status), data=twindata)
summary(twinreg)
```
```{r}
twinreg2<-lm(IQF~IQN+status+IQN:status, data=twindata)
summary(twinreg2)
```
```{r}
twinreg2<-lm(IQF~IQN+status+IQN:status, data=twindata)
summary(twinreg2)
```
```{r}
anova(twinreg, twinreg2)
```

```{r}
twindata$baseH<-relevel(twindata$status, ref="h")
twindata$baseM<-relevel(twindata$status, ref="m")
twindata$baseL<-relevel(twindata$status, ref="l")
as.factor(twindata$status)
```
```{r}
twindata$baseH
```

```{r}
twindata$baseM

```

```{r}
twindata$baseL
```
```{r}
twinreg.baseH <- lm(IQF~IQN+baseH, data=twindata)
twinreg.baseM <- lm(IQF~IQN+baseM, data=twindata)
twinreg.baseL <- lm(IQF~IQN+baseL, data=twindata)
```

#2.4 Interpret the regression coefficients in the context of the problem.
the regression suggest that the child that was natured by a natural parent had a high IQ score compared to the other child.